Storm Surges

### 2014-04-24 RoughWork/manuscript.tex
Fixed typos

\subsection{Model domain}
The modelled domain extends from the Strait of Juan de Fuca to Puget Sound to Johnstone Strait as shown in Figure \ref{fig:domain}. Bathymetry from the Cascadia physiography dataset \citep{haugerud1999digital} was smoothed to limit the difference in depth across grid cells. For model stability, additional smoothing at the Strait of Juan de Fuca western boundary was imposed to achieve constant depth across the first ten grid cells. As depicted in Figure \ref{fig:domain}, the numerical grid is rotated $29^{\circ}$ counter-clockwise of North in order to maintain computational efficency since currents within the Strait of Georgia are mainly aligned with this rotated axis.

The curvilinear orthogonal numerical grid is divided into 400 by 900 by 40 grid cells, which results in an almost uniform horizontal resolution with grid spacing approximately 400 m by 500 m. The 40 vertical $z$-levels are stretched gradually in order to achieve higher resolution in the surface layer, with 1 m vertical grid spacing down to about 10 m in depth. Below 10m the grid is stretched gradually with a maxmimum grid spacing of 27 m at the lowest layer. At the bottom boundary, partial $z$-levels are utilized in order to limit large changes in bathymetry across grid cells \citep{madec2008nemo}.

@@

\subsection{Boundary conditions and subgrid scale processes}
%lateral and bottom friction, open boundaries, eta surface
The model includes two open boundaries that connect to the Pacific Ocean, the western boundary of the Strait of Juan de Fuca as well as Johnstone Strait at the north, both of which are forced with eight tidal constituents, temperature and salinity climatologies, and the sea surface height anomaly. The details of these forcing conditions are described below. At coastal boundaries, the partial slip boundary condition, an approximation of the no slip boundary condition, is used. The partial slip boundary condition allows one to include the frictional effects of lateral boundaries without the restricitve resolution required to represent the lateral boundary layer under no slip conditions. A lateral eddy viscosity of 20 m $^2$ s $^{-1}$ parameterizes horizontal friction and the lateral eddy diffusivity is 20.5 m $^2$ s $^{-1}$ . At the ocean surface, meterological conditions and fresh water input from rivers are included and described in more detail below. Bottom friction is represented by a quadratic law for the bottom momentum flux with drag coefficient $C_D = 5\times 10^{-3}$. Vertical turbulence and mixing is calculated through the $k-\epsilon$ configuration of the generic length scale (GLS) turbulence closure \citep{umlauf2003generic} with background vertical eddy viscosity and diffusivity set to $1\times10^{-4}$ and $1\times10^{-5}$ m$^2$ s$^{-1}$ respectively. Details on the NEMO implementation of the partial slip lateral boundary condition, quadratic bottom friction law, and GLS turbulence closure scheme are provided by \citep{madec2008nemo}.

@@

\subsubsection{Tidal forcing} 
The model was forced by tidal elevations and currents at the Juan de Fuca and Johnstone Strait boundaries. Tidal heights and currents for eight tidal constituents (K1 ,O1, P1, Q1, M2, K2, N2, S2) at grid points along the Juan de Fuca boundary were extracted from Webtide, an online web prediction model for the northeast Pacific Ocean, which is based on work by \citet{foreman2000webtide}. The Johnstone Strait boundary was forced with current and elevation tidal harmonics measured and calculated by \citet{thomson1980johnstone} for the major M2 and K1 constituents. Additionally, O1 and S2 elevation harmonics from their measuresments were employed. The remaining consituents were extrapolated from Webtide.

@@

\subsubsection{Sea Surface Height}
Sea surface height at the mouth of Juan de Fuca was set using values from the Tofino tide guage.  A monthly climatology was produced using daily averages from 2000-2010, binning them by month, averaging and setting the yearly mean to zero.  For the storm surge simulations, hourly variations in sea surface height were used.  These values are the Tofino tide guage values, de-tided and with the zero reset as for the climatology.

@@

The model relaxed to the forced temperature and salinity over the 10 grid points (about 5km) closest to the open boundaries, using the NEMO FRS scheme \citep{madec2008nemo}. %other references>
The tidal forcing and sea surface height was used in the barotropic velocity forcing which used the NEMO Flather scheme \citet{madec2008nemo}. %other references?
The baroclinic velocities at the boundary were set to be equal to the values inside the boundary (zero-gradient boundary conditions).  This scheme is not part of core NEMO.  Zero gradient conditions were chosen because the baroclinic velocity at the mouth of Juan de Fuca is primarily estuarine and thus set by density variations between inside and outside the domain.

@@

The monthly flow from each river was input as a point source in the three grid points closest to the surface at the model point closest to the mouth of each river. Incoming water was assumed to be fresh and at surface temperatyre. A total of 150 rivers were parameterised by this method.

@@

\subsection{Initial conditions}
Initial conditions for temperature and salinity were taken from a CTD cast in the middle Strait of Georgia taken in Sept 2002 \citep{pawlowiczetal2007}.  Conditions were initially uniform horizontally.  Velocity was initialized at zero. 
%Merge with below?
%Perhaps image of spun-up stratifiction along Thalweg? Here or in model evaluation

@@

\section{Storm Surge Hindcasts}\label{sec:storm}

>>>>

\subsection{Model domain}
The modelled domain extends from the Strait of Juan de Fuca to Puget Sound to Johnstone Strait as shown in Figure \ref{fig:domain}. Bathymetry from the Cascadia physiography dataset \citep{haugerud1999digital} was smoothed to limit the difference in depth across grid cells. For model stability, additional smoothing at the Strait of Juan de Fuca western boundary was imposed to achieve constant depth across the first ten grid cells. As depicted in Figure \ref{fig:domain}, the numerical grid is rotated $29^{\circ}$ counter-clockwise of North in order to maintain computational efficiency since currents within the Strait of Georgia are mainly aligned with this rotated axis.

The curvilinear orthogonal numerical grid is divided into 400 by 900 by 40 grid cells, which results in an almost uniform horizontal resolution with grid spacing approximately 400 m by 500 m. The 40 vertical $z$-levels are stretched gradually in order to achieve higher resolution in the surface layer, with 1 m vertical grid spacing down to about 10 m in depth. Below 10m the grid is stretched gradually with a maximum grid spacing of 27 m at the lowest layer. At the bottom boundary, partial $z$-levels are utilized in order to limit large changes in bathymetry across grid cells \citep{madec2008nemo}.

@@

\subsection{Boundary conditions and sub-grid scale processes}
%lateral and bottom friction, open boundaries, eta surface
The model includes two open boundaries that connect to the Pacific Ocean, the western boundary of the Strait of Juan de Fuca as well as Johnstone Strait at the north, both of which are forced with eight tidal constituents, temperature and salinity climatologies, and the sea surface height anomaly. The details of these forcing conditions are described below. At coastal boundaries, the partial slip boundary condition, an approximation of the no slip boundary condition, is used. The partial slip boundary condition allows one to include the frictional effects of lateral boundaries without the restrictive resolution required to represent the lateral boundary layer under no slip conditions. A lateral eddy viscosity of 20 m $^2$ s $^{-1}$ parametrizes horizontal friction and the lateral eddy diffusivity is 20.5 m $^2$ s $^{-1}$ . At the ocean surface, meteorological conditions and fresh water input from rivers are included and described in more detail below. Bottom friction is represented by a quadratic law for the bottom momentum flux with drag coefficient $C_D = 5\times 10^{-3}$. Vertical turbulence and mixing is calculated through the $k-\epsilon$ configuration of the generic length scale (GLS) turbulence closure \citep{umlauf2003generic} with background vertical eddy viscosity and diffusivity set to $1\times10^{-4}$ and $1\times10^{-5}$ m$^2$ s$^{-1}$ respectively. Details on the NEMO implementation of the partial slip lateral boundary condition, quadratic bottom friction law, and GLS turbulence closure scheme are provided by \citep{madec2008nemo}.

@@

\subsubsection{Tidal forcing} 
The model was forced by tidal elevations and currents at the Juan de Fuca and Johnstone Strait boundaries. Tidal heights and currents for eight tidal constituents (K1 ,O1, P1, Q1, M2, K2, N2, S2) at grid points along the Juan de Fuca boundary were extracted from Webtide, an online web prediction model for the northeast Pacific Ocean, which is based on work by \citet{foreman2000webtide}. The Johnstone Strait boundary was forced with current and elevation tidal harmonics measured and calculated by \citet{thomson1980johnstone} for the major M2 and K1 constituents. Additionally, O1 and S2 elevation harmonics from their measurements were employed. The remaining constituents were extrapolated from Webtide.

@@

\subsubsection{Sea Surface Height}
Sea surface height at the mouth of Juan de Fuca was set using values from the Tofino tide gauge.  A monthly climatology was produced using daily averages from 2000-2010, binning them by month, averaging and setting the yearly mean to zero.  For the storm surge simulations, hourly variations in sea surface height were used.  These values are the Tofino tide gauge values, de-tided and with the zero reset as for the climatology.

@@

The model relaxed to the forced temperature and salinity over the 10 grid points (about 5km) closest to the open boundaries, using the NEMO flow relaxation scheme \citep{madec2008nemo}. %other references>
The tidal forcing and sea surface height was used in the barotropic velocity forcing which used the NEMO Flather scheme \citep{madec2008nemo}. %other references?
The baroclinic velocities at the boundary were set to be equal to the values inside the boundary (zero-gradient boundary conditions).  This scheme is not part of core NEMO.  Zero gradient conditions were chosen because the baroclinic velocity at the mouth of Juan de Fuca is primarily estuarine and thus set by density variations between inside and outside the domain.

@@

The monthly flow from each river was input as a point source in the three grid points closest to the surface at the model point closest to the mouth of each river. Incoming water was assumed to be fresh and at surface temperature. A total of 150 rivers were parameterised by this method.

@@

\subsection{Initial conditions}
Initial conditions for temperature and salinity were taken from a CTD cast in the middle Strait of Georgia taken in Sept 2002 \citep{pawlowiczetal2007}.  Conditions were initially uniform horizontally.  Velocity was initialized at zero. 
%Merge with below?
%Perhaps image of spun-up stratification along Thalweg? Here or in model evaluation

@@

\section{Storm Surge Hind casts}\label{sec:storm}
